---
title: "GroupE_Classification"
date: "5/29/2019"
output:
  html_document:
    toc: true
    toc_depth: 3
author: "Group E"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

start_time <- Sys.time()
print(paste0('---START--- Starting at ',start_time))

packages_list <- c('ggplot2',
                   'data.table'
                    )

for (i in packages_list){
  if(!i%in%installed.packages()){
    install.packages(i, dependencies = TRUE)
    library(i, character.only = TRUE)
    print(paste0(i, ' has been installed'))
  } else {
    print(paste0(i, ' is already installed'))
    library(i, character.only = TRUE)
  }
}

print(paste0('[', round(difftime(Sys.time(),start_time, units = 'secs'),1), 's]: ',
             'All necessary packages installed and loaded'))
```

# Classifying Term Deposits

## 1. Introduction

## 2. Data Loading and Preprocessing 

## 3. Data Exploration

## 4. Baseline

## 5. Feature Engineering

## 6. Modeling

## 1. Introduction
This is the intro.

## 2. Data Loading and Preprocessing

Here we load the necessary train and test data.

```{r Load Data, include=FALSE}
# Loading the Data
train = fread("BankCamp_train.csv", stringsAsFactors = T)
test = fread("BankCamp_test.csv", stringsAsFactors = T)
```

These are the first few rows of the training data.

```{r head}
head(train)
```

This is the structure of the data.

```{r str}
str(train)
```

This is the summary of the data.

```{r summary}
summary(train)
```

Now we check if there are missing values in any column:

```{r missing}
sapply(train, function(x) sum(is.na(x)))
```

Finally we check if there are any duplicate values:

```{r duplicated}
any(duplicated(train))
```

Now we set the types for numerical variables:

```{r numerical}
train[ , which(sapply(train, is.integer)):=lapply(.SD,as.numeric), .SDcols = sapply(train, is.integer)]
test[ , which(sapply(test, is.integer)):=lapply(.SD,as.numeric), .SDcols = sapply(test, is.integer)]
```

